This dissertation marks the end of a long journey that certainly was not possible on my own.
I would first like to thank my Ph.D.\ advisors, Prof.\ Kathryn D.\ Huff and Prof.\ Madicken Munk. I
am deeply grateful to Prof.\ Huff for her technical guidance on multiphysics reactor modeling \&
simulation and best practices in software development. I am equally grateful to Prof.\ Munk for her
guidance on neutron transport methods for
this dissertation. I am also grateful to both advisors for their support and advice on time
management, scientific writing, and other matters. In many ways, they created a welcoming
and supportive environment in the Advanced Reactors and Fuel Cycles (ARFC)
research group, without which this dissertation would not have been possible. I would also like to
thank my Ph.D.\ committee members, Prof.\ Rizwan Uddin, Prof.\ Tomasz Kozlowski, and Prof.\ Paul
Fischer, for teaching me much of what I know today about reactor physics and computational fluid
dynamics and for their guidance on this dissertation.

I am thankful to current and former groupmates in ARFC---Sam Dotson, Nathan Ryan, Luke Seifert,
Olek Yardas, Zoe Richter, Nathan Glaser, Gwendolyn Chee, Amanda Bachmann,
Andrei Rykhlevskii, Anshuman Chaube, and Greg Westphal---for their friendship and their help with
code and writing reviews. I am also thankful to
Alvin Lee, Xin Zhi Tan, Jeremy Mettler, and other friends from NPRE and UIUC
for the fun times we shared and for their help whenever I needed it.

I am especially thankful to my partner, Gwendolyn Chee, for her constant support and encouragement,
which spurred me on through the toughest times.
Finally, I want to thank my family for their unwavering love and support. I would not be where I am
today without them.

I would also like to acknowledge funding support for my graduate studies from the Singapore
Nuclear Safety \& Research Initiative (SNRSI) and the Department of Nuclear, Plasma \& Radiological
Engineering (NPRE) at the University of Illinois Urbana-Champaign (UIUC).

This research used resources of the Argonne Leadership Computing Facility, a U.S.\ Department of
Energy (DOE) Office of Science user facility at Argonne National Laboratory, and is based on
research supported by the U.S.\ DOE Office of Science---Advanced Scientific Computing Research
Program, under Contract No.\ DE-AC02-06CH11357.
