Section \ref{sec:msr-vv} identified salt flow-induced \gls{DNP} drift and the strong coupling
between neutronics and thermal-hydraulics as two critical multiphysics interactions that must be
accurately modeled in \gls{MSR} simulation software. This chapter describes Moltres and
subsequently presents several \gls{VV} results and analyses for verifying and validating specific
modeling capabilities in Moltres which relate to the aforementioned multiphysics interactions.
Section \ref{sec:moltres-description} details the neutronics, thermal-hydraulics, and multiphysics
coupling schemes in Moltres for \gls{MSR} multiphysics simulations. Following the model
descriptions, the section also summarizes and assesses previous \gls{MSR} analyses
performed with Moltres. Section \ref{sec:cnrs-benchmark} presents verification results from Moltres
for the CNRS benchmark study. Section \ref{sec:msre-pump} presents key findings from a
collaborative \gls{VV} study based on the \gls{MSRE} pump transient tests conducted in the 1960s.
Section \ref{sec:turbulence} describes the \gls{SA} turbulence model implementation in
Moltres and presents the model \gls{VV} results against common wall-bounded turbulent flow case
studies from literature.

\input{moltres}
\input{benchmark}

\input{msre-pump}

\input{turbulence}

\FloatBarrier

\section{Summary}

Moltres is a highly scalable \gls{MOOSE}-based multiphysics application for advanced reactor
simulations with a strong focus on \glspl{MSR} modeling. It features the multigroup neutron
diffusion and thermal-hydraulics models, along with robust and flexible coupling capabilities for
modeling multiphysics interactions in reactors.

This chapter presented two \gls{VV} studies aimed at primarily assessing Moltres for modeling
two key multiphysics interactions in \glspl{MSR}, namely looped \gls{DNP} flow modeling and the
strong coupling between neutronics and thermal-hydraulics. The first study is the CNRS numerical
benchmark for fast-spectrum \gls{MSR} modeling. The simulation results showed that Moltres is
consistent with the four \gls{MSR} simulation software that participated in the CNRS benchmark
study \cite{tiberga_results_2020} over a wide range of neutronics and thermal-hydraulics simulation
metrics. Moltres underpredicts the temperature along the top boundary due to unphysical,
discontinuous velocity boundary conditions defined at the top corners of the lid-driven cavity
flow setup. Nevertheless, the temperature deviations are limited to the top boundary and Moltres
is otherwise in excellent agreement with the benchmark participants.

The second study is a collaborative work with the developer of QuasiMolto for assessing \gls{DNP}
flow modeling in the \gls{MSRE} zero-power pump start-up and coast-down transient tests conducted
in the 1960s \cite{prince_zero-power_1968}. Despite the widely different numerical schemes between
Moltres and QuasiMolto, both codes were consistent with each other under all considered scenarios.
$k_\text{eff}$ estimates from both codes were 22 pcm and 21 pcm apart under static and steady salt
flow conditions. Consequently, the calculated reactivity loss due to \gls{DNP} flow agreed within
less than 1 pcm. Spatial flux and \gls{DNP} distributions showed nearly perfect overlap with less
than 0.8\% relative difference between Moltres and QuasiMolto. Moltres and QuasiMolto remained
consistent with each other in the pump start-up and coast-down transients. We attributed
discrepancies between the numerical results and \gls{ORNL} experimental data to the absence of
upper and lower plena in the numerical model. The numerical study remains a useful benchmark for
assessing \gls{DNP} advection modeling in \gls{MSR} simulation software due to clearly defined
salt recirculation behavior reflected in the results.

Lastly, this chapter presented the implementation and verification of the \gls{SA} model
\cite{spalart_one-equation_1994} in Moltres for enabling wall-bounded turbulent molten salt flow
modeling. The \gls{SA} model includes a rotation correction scheme
\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} for improved accuracy when
modeling turbulent flow with curved streamlines. The \gls{SA} model performed well under the
turbulent channel flow verification test \cite{moser_direct_1999}, the turbulent pipe flow
validation test \cite{laufer_structure_1954}, and the \gls{BFS} flow validation test
\cite{driver_features_1985}.
